Self Referencing Sequences

### 2014-11-05 draft.tex
changed sequences to curly braces

\begin{lemma}
Let $C$ be a generating cycle. Suppose $(a_i)$ and $(b_i)$ are both self-referencing sequences with $a_j \neq b_j$ and $j$ is the smallest such number, then $j = 0$.
\end{lemma}

\begin{proof}
Suppose $j > 0$, then $a_{j-1} = b_{j-1}$ since $j$ was assumed to be the smallest index where $(a_i)$ and $(b_i)$ are different. Suppose $a_{j-1} \neq a_j$ and $b_{j-1} \neq b_j$, then we can ask what comes directly after $a_{j-1}$ is the generating cycle; by looking at sequence $(a_i)$, we get that $a_j$ comes after $a_{j-1}$, but by looking at the sequence $(b_i)$, we get that $b_j$ comes after $a_{j-1}$. This is a contradiction since $a_j \neq b_j$. So lets assume without loss of generality that $a_{j-1} = a_j$, and also, we will assume that $a_{j-1}$ belongs to the $k$th block. Since the sequences are the same up to index $j$, $b_{j-1}$ belongs to the $k$th block. However, this means that $a_k \geq b_k + 1$. So $k \geq j$ and $k$ designates the block length of the block containing $a_j$, and so $k\leq j$, so $k=j$. Figure~\ref{fig:unequalsequences} gives a visual representation of these sequences.

@@

We now introduce a notion which will help us categorize sequences as equivalent. 
\begin{definition} Consider a sequence $(a_i)$. We define the read distance for an index $i$ as the number of terms in between it $a_i$ and the beginning of the $i$th block of numbers.
\end{definition}
The read distance can be any nonnegative integer. For example if we consider the sequence $(a_i)$ generated by $(1,2,3)$ beginning with a 1  
\begin{align*}
1,2,2,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
\end{align*}
the read distance for $1$ is always $0$ since in a self-referencing sequence, the first number must refer to the first block. The read distance for $2$ is $0$ as the second block of numbers starts at the second number. We can continue to calculate the read distances for this sequence and see that they form a monotonically increasing sequence $0, 0 ,1,2,4,6,6,6,6,7,8,9,\ldots$. The read distance, can be thought of as the horizontal shift of the arrows in Figure~\ref{fig:onetwothree}.

\begin{lemma} 
\label{readdistformula}
Let $(a_i)$ be any self referencing sequence generated by some set of positive integers. Then the read distance for $i>1$, is 
\begin{equation*}
\sum_{k=1}^{i-1}( a_k - 1)
\end{equation*} 
and the read distance for $1$ is always 0.
\end{lemma}
\begin{proof}
We will prove this using induction. We know that for $i=1$ the read distance is $0$, as the first block always starts at the first number. When $i=2$ the read distance is $a_1-1$. The first block will be of length $a_1$ since the sequence is self referencing, therefore the second block will begin at position $a_1$. The number of elements between position $a_1$ and $2$ is exactly $a_1 -1$.

@@

\begin{lemma} Consider a self referencing sequence $(a_i)$ generated by $(c_1, c_2, \ldots, c_n)$ with $m = \max(c_1, c_2, \ldots, c_n)$. Suppose the read distance for $i$ is $d$, then the read distance for $i+m$ is at least $d+1$. 
\end{lemma}
\begin{proof}
The read distances can stay constant from $i$ to $i+1$ only if there is a $a_{i+1} = 1$. There can be at most $m$ 1's in a contiguous block, so the read distance of $i+m$ is bigger than the read distance of $i$.  
\end{proof}

@@

We also have a sufficient condition for equivalence 
\begin{proposition} 
\label{equivsuff}
If two sequences $(a_i), (b_i)$ are such that for some integers $m,l$ the read distance for $a_m$ and $b_l$ are both $d$, some integer and 
\begin{equation*}
a_{m+k} = b_{l+k}
\end{equation*}
where $0 \leq k \leq d-1$ then $\{a_i\}$ and $\{b_i\}$ are equivalent. 
\end{proposition}
\begin{proof}
We will prove this by  induction. In particular we will show for all $k \geq 0$ that $a_{m+k} = b_{l+k}$ and the read distance for $a_{m+k}$ equals the read distance for $b_{l+k}$. By our assumptions we already have this for $0 \leq k \leq d-1$ and so we begin with $k=d$.

@@

The following is a simple observation of sequences that begin with $1$.

\begin{corollary} Let $C = (1 c_1 c_2 \cdots c_n)$. $s(C, 1)$ and $s(C, c_1)$ are equivalent.
\end{corollary}
\begin{proof}
Let $\{a_i\} = s(C, 1)$ and $\{b_i\} = s(C, c_1)$. The read distance of sequence $\{a_i\}$ for $i=2$ is 0, and the read distance of sequence $\{b_i\}$ for $i = 1$ is 0. In addition, $a_2 = b_1 = c_1$, so by Proposition ~\ref{equivsuff} the two sequences are equivalent.
\end{proof}

@@

\begin{conjecture} If two sequences $(a_i)$ and $(b_i)$ have elements $a_k \neq b_l$, for $k,l>1$ with the same read distance then they are not equivalent. 
\end{conjecture}

>>>>

\begin{lemma}
Let $C$ be a generating cycle. Suppose $\{a_i\}$ and $\{b_i\}$ are both self-referencing sequences with $a_j \neq b_j$ and $j$ is the smallest such number, then $j = 0$.
\end{lemma}

\begin{proof}
Suppose $j > 0$, then $a_{j-1} = b_{j-1}$ since $j$ was assumed to be the smallest index where $\{a_i\}$ and $\{b_i\}$ are different. Suppose $a_{j-1} \neq a_j$ and $b_{j-1} \neq b_j$, then we can ask what comes directly after $a_{j-1}$ is the generating cycle; by looking at sequence $\{a_i\}$, we get that $a_j$ comes after $a_{j-1}$, but by looking at the sequence $\{b_i\}$, we get that $b_j$ comes after $a_{j-1}$. This is a contradiction since $a_j \neq b_j$. So lets assume without loss of generality that $a_{j-1} = a_j$, and also, we will assume that $a_{j-1}$ belongs to the $k$th block. Since the sequences are the same up to index $j$, $b_{j-1}$ belongs to the $k$th block. However, this means that $a_k \geq b_k + 1$. So $k \geq j$ and $k$ designates the block length of the block containing $a_j$, and so $k\leq j$, so $k=j$. Figure~\ref{fig:unequalsequences} gives a visual representation of these sequences.

@@

We now introduce a notion which will help us categorize sequences as equivalent. 
\begin{definition} Consider a sequence $\{a_i\}$. We define the read distance for an index $i$ as the number of terms in between it $a_i$ and the beginning of the $i$th block of numbers.
\end{definition}
The read distance can be any nonnegative integer. For example if we consider the sequence $\{a_i\}$ generated by $(1,2,3)$ beginning with a 1  
\begin{align*}
1,2,2,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
\end{align*}
the read distance for $1$ is always $0$ since in a self-referencing sequence, the first number must refer to the first block. The read distance for $2$ is $0$ as the second block of numbers starts at the second number. We can continue to calculate the read distances for this sequence and see that they form a monotonically increasing sequence $0, 0 ,1,2,4,6,6,6,6,7,8,9,\ldots$. The read distance, can be thought of as the horizontal shift of the arrows in Figure~\ref{fig:onetwothree}.

\begin{lemma} 
\label{readdistformula}
Let $\{a_i\}$ be any self referencing sequence generated by some set of positive integers. Then the read distance for $i>1$, is 
\begin{equation*}
\sum_{k=1}^{i-1}( a_k - 1)
\end{equation*} 
and the read distance for $1$ is always 0.
\end{lemma}
\begin{proof}
We will prove this using induction. We know that for $i=1$ the read distance is $0$, as the first block always starts at the first number. When $i=2$ the read distance is $a_1-1$. The first block will be of length $a_1$ since the sequence is self referencing, therefore the second block will begin at position $a_1$. The number of elements between position $a_1$ and $2$ is exactly $a_1 -1$.

@@

\begin{lemma} Consider a self referencing sequence $\{a_i\}$ generated by $(c_1, c_2, \ldots, c_n)$ with $m = \max(c_1, c_2, \ldots, c_n)$. Suppose the read distance for $i$ is $d$, then the read distance for $i+m$ is at least $d+1$. 
\end{lemma}
\begin{proof}
The read distances can stay constant from $i$ to $i+1$ only if there is a $a_{i+1} = 1$. There can be at most $m$ 1's in a contiguous block, so the read distance of $i+m$ is bigger than the read distance of $i$.  
\end{proof}

@@

We also have a sufficient condition for equivalence 
\begin{proposition} 
\label{equivsuff}
If two sequences $\{a_i\}, \{b_i\}$ are such that for some integers $m,l$ the read distance for $a_m$ and $b_l$ are both $d$, some integer and 
\begin{equation*}
a_{m+k} = b_{l+k}
\end{equation*}
where $0 \leq k \leq d-1$ then $\{a_i\}$ and $\{b_i\}$ are equivalent. 
\end{proposition}
\begin{proof}
We will prove this by  induction. In particular we will show for all $k \geq 0$ that $a_{m+k} = b_{l+k}$ and the read distance for $a_{m+k}$ equals the read distance for $b_{l+k}$. By our assumptions we already have this for $0 \leq k \leq d-1$ and so we begin with $k=d$.

@@

This result allows us to make a general statement for equivalence for sequences that begin with a $1$.
\begin{corollary} If $C = (1 c_1 c_2 \cdots c_n)$, then $s(C, 1)$ and $s(C, c_1)$ are equivalent.
\end{corollary}
\begin{proof}
Let $\{a_i\} = s(C, 1)$ and $\{b_i\} = s(C, c_1)$. The read distance of sequence $\{a_i\}$ for $i=2$ is 0, and the read distance of sequence $\{b_i\}$ for $i = 1$ is 0. In addition, $a_2 = b_1 = c_1$, so by Proposition ~\ref{equivsuff} the two sequences are equivalent.
\end{proof}

@@

\begin{conjecture} If two sequences $\{a_i\}$ and $\{b_i\}$ have elements $a_k \neq b_l$, for $k,l>1$ with the same read distance then they are not equivalent. 
\end{conjecture}
